Notes

### 2022-09-04 chapters/introduction.tex
Finish first complete revision

In recent years numerous proposals have arisen on how to suitably deal with functional data, all of them encompassed under the term Functional Data Analysis (FDA), which essentially explores statistical techniques to process, model and make inference on data varying over a continuum. A partial survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{ramsay2005functional}, \citet{hsing2015theoretical} or the book by \citet{horvath2012inference}. As the name suggests, FDA techniques are heavily inspired by functional analysis tools and methods: Hilbert spaces, orthonormal systems, linear operators, and so on. In particular, a notion that also intersects with the classical theory of machine learning and pattern recognition is that of reproducing kernel Hilbert spaces (RKHS's). We will demonstrate throughout this work how these spaces of functions possess properties that allow for an efficient treatment of functional data.

@@

In this work we are concerned with functional linear and logistic regression models, that is, situations where the goal is to predict a continuous or dichotomous variable from functional observations. Even though these problems can be formally stated with almost no differences from their finite-dimensional counterparts, there are some fundamental challenges as well as some subtle drawbacks that emerge as a result of working in infinite dimensions. Moreover, we will concentrate our efforts on the case in which the response is a scalar, though function-on-function and function-on-scalar regression are also interesting scenarios amply explored in the literature.

To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\) (i.e. a process with finite second moments and whose realizations are square-integrable functions indexed on \([0,1]\)). The underlying probability space \((\Omega, \mathcal A, \P)\) is not important. We will further suppose without loss of generality that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\). In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\D_n =\{(x_i, y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), where the functional observations are recorded on a common finite grid \(\{t_j\}\subset [0, 1]\). Our ultimate aim will be to accurately predict the response corresponding to unlabeled samples from \(X\). Figure~\ref{fig:linear_data_example} depicts a typical data set used in functional regression, and we already saw in Figure~\ref{fig:tecator} what a functional classification data set may look like.

\begin{figure}[ht]
  \centering
  \includegraphics[width=0.9\textwidth]{linear_data_example}
  \caption{Simulated data set for functional regression. On the left we have \(n=50\) functional observations on an equispaced grid of \(N=100\) points on \([0,1]\). To each observation corresponds a real number; the distribution of these responses is shown on the right.}\label{fig:linear_data_example}
\end{figure}

>>>>

In recent years numerous proposals have arisen on how to suitably deal with functional data, all of them encompassed under the term Functional Data Analysis (FDA), which essentially explores statistical techniques to process, model and make inference on data varying over a continuum. A partial survey on such techniques and methods is \citet{cuevas2014partial}, while a more detailed exposition of the theory and applications can be found for example in \citet{ramsay2005functional}, \citet{hsing2015theoretical} or the book by \citet{horvath2012inference}. As the name suggests, FDA techniques are heavily inspired by functional analysis tools and methods: Hilbert spaces, orthonormal systems, linear operators, and so on. In particular, a notion that also intersects with the classical theory of machine learning and pattern recognition, and that has gained traction in recent years, is that of reproducing kernel Hilbert spaces (RKHS's). We will demonstrate throughout this work how these spaces of functions possess properties that allow for an efficient treatment of functional data.

@@

In this work we are concerned with functional linear and logistic regression models, that is, situations where the goal is to predict a continuous or dichotomous variable from functional observations. Even though these problems can be formally stated with almost no differences from their finite-dimensional counterparts, there are some fundamental challenges as well as some subtle drawbacks that emerge as a result of working in infinite dimensions. Moreover, we will concentrate our efforts on the case in which the response is a scalar, though function-on-function and function-on-scalar regression are also interesting scenarios widely explored in the literature. To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\) (i.e. a process with finite second moments and whose realizations are square-integrable functions indexed on \([0,1]\)). The underlying probability space \((\Omega, \mathcal A, \P)\) is not important. We will further suppose that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\). In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\D_n =\{(x_i, y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), where the functional observations are recorded on a common finite grid \(\{t_j\}\subset [0, 1]\). Our ultimate aim will be to accurately predict the response corresponding to unlabeled samples from \(X\). Figure~\ref{fig:linear_data_example} depicts a typical data set used in functional regression, and we already saw in Figure~\ref{fig:tecator} what a functional classification data set may look like.

\enlargethispage{1\baselineskip}

\begin{figure}[htbp!]
  \centering
  \includegraphics[width=0.9\textwidth]{linear_data_example}
  \caption{Simulated data set for functional regression. On the left we have \(n=50\) functional observations on an equispaced grid of \(N=100\) points on \([0,1]\). To each observation corresponds a real number; the distribution of these responses is shown on the right.}\label{fig:linear_data_example}
\end{figure}

### 2022-09-04 chapters/model-choice.tex
Finish first complete revision

The most straightforward solutions consist on imposing an artificial identifiability constraint on the parameters to break the symmetry of their posterior distributions; see \citet{jasra2005markov} and references therein. A common approach that seems to work well is to simply impose an ordering in the offending parameters, which in our case would mean requiring for example that \(\beta_i < \beta_j\) for \(i < j\), or the analogous with the \(\tau_j\). We have implemented a variation of this method described in \citet{simola2021approximate}, which works by post-processing the samples and relabeling the components to satisfy the order constraint mentioned above, choosing either \(b\) or \(\tau\) depending on which set of ordered parameters would produce the largest separation between any two of them (suitably averaged across all iterations of the chains).

>>>>

The most straightforward solutions consist on imposing an artificial identifiability constraint on the parameters to break the symmetry of their posterior distributions; see \citet{jasra2005markov} and references therein. A common approach that seems to work well is to simply impose an ordering in the parameters in question, which in our case would mean requiring for example that \(\beta_i < \beta_j\) for \(i < j\), or the analogous with the \(\tau_j\). We have implemented a variation of this method described in \citet{simola2021approximate}, which works by post-processing the samples and relabeling the components to satisfy the order constraint mentioned above, choosing either \(b\) or \(\tau\) depending on which set of ordered parameters would produce the largest separation between any two of them (suitably averaged across all iterations of the chains).
